\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{microtype}
\usepackage{tikz}
\usepackage{float}
\usetikzlibrary{positioning,arrows.meta}

\setstretch{1.08}
\setlist[itemize]{leftmargin=1.5em}
\setlist[enumerate]{leftmargin=1.8em}

\title{\textbf{Style Forecast: Technical Report}}
\author{COMP3011 Coursework 1}
\date{\today}

\begin{document}
\maketitle

\section{Overview}
Style Forecast is an API-first outfit recommendation system designed for a practical everyday problem: people often have limited time when getting ready, incomplete awareness of weather conditions at a chosen time, and only partial knowledge of current style trends. The system aims to bridge that knowledge gap by combining weather, trend, and user-preference signals into a recommendation that helps the user feel both comfortable and confident.

The central design objective was not simply to return an outfit, but to return one that is appropriate for climate, relevant to current fashion context, and explainable enough to earn user trust.

\section{Methodology}
\subsection{Problem Framing and Design Approach}
The project was approached as a decision-support system rather than a generic fashion catalogue. That distinction shaped the methodology. A recommendation is only useful if it reduces uncertainty at the point of decision-making, so the model needs to reason about context, not only preference. For that reason, the chosen methodology combines:
\begin{itemize}
  \item \textbf{weather suitability}, to reduce the risk of physical discomfort,
  \item \textbf{trend relevance}, to reflect recent fashion signals,
  \item \textbf{user feedback adaptation}, to avoid static, one-size-fits-all behaviour.
\end{itemize}

I considered alternative approaches before selecting the final model. Pure collaborative filtering was not suitable because the dataset is small and suffers from cold-start problems. An end-to-end LLM ranking approach was also rejected because it would be harder to audit and less reliable for a live assessment setting. I therefore chose a hybrid weighted model that exposes its reasoning while still incorporating multiple data sources.

\subsection{Architecture Overview}
The system is structured in three layers, shown in Figure~\ref{fig:architecture}. This separation keeps the application maintainable and makes design choices easier to justify.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  node distance=5mm and 6mm,
  every node/.style={font=\footnotesize},
  box/.style={draw, rounded corners, align=center, text width=0.31\linewidth, minimum height=0.82cm, inner sep=3pt},
  smallbox/.style={draw, rounded corners, align=center, text width=0.24\linewidth, minimum height=0.82cm, inner sep=3pt},
  arr/.style={-{Latex[length=2mm]}, thick}
]
\node[box] (client) {React + Vite Client\\user input, authentication state, and visual feedback};
\node[box, below=of client] (api) {NestJS API\\auth, items, outfits, climate, trends, recommendation};
\node[smallbox, below left=of api] (db) {PostgreSQL + Prisma\\relational persistence};
\node[smallbox, below right=of api] (ext) {External sources\\Open-Meteo and trend feeds};
\draw[arr] (client) -- (api);
\draw[arr] (api) -- (db);
\draw[arr] (ext) -- (api);
\end{tikzpicture}
\caption{High-level architecture of Style Forecast.}
\label{fig:architecture}
\end{figure}

The client is intentionally thin: it collects user input, presents results, and manages authenticated state. The API is the coordination layer for validation, scoring, ownership rules, and external data access.

\subsection{Technology Choices and Rationale}
\textbf{TypeScript} was chosen across frontend and backend to reduce integration mismatch and improve refactor safety \cite{typescript}. This was a practical choice rather than a stylistic one: using one typed language simplified DTO design, API contracts, and debugging.

\textbf{NestJS} was chosen for the backend because its modular structure fits a system with distinct domains such as authentication, items, outfits, climate, trends, and recommendations \cite{nestjs}. Keeping logic in services rather than controllers improved clarity and made it easier to adjust scoring logic without destabilising routing behaviour.

\textbf{PostgreSQL} with \textbf{Prisma} was chosen for persistence \cite{postgresql,prisma}. An SQL database is appropriate because the project is relational: users own items, outfits contain items, feedback references users and targets, and time-based records such as climate snapshots and trend entries benefit from structured querying. Prisma added migrations and type-safe data access, which reduced implementation errors.

\textbf{React + Vite} was chosen for the client because the interaction model is stateful and iterative \cite{react,vite}. React supports interactive recommendation flows and authenticated UI state, while Vite keeps iteration fast.

\textbf{REST over GraphQL} was also a deliberate choice. GraphQL would have been defensible for highly variable client-side queries, but this project has a relatively stable set of resources and workflows. REST made route semantics, status codes, authentication guards, and Swagger-based documentation more direct \cite{openapi}.

Another important design choice was to keep modules narrow in scope. For example, climate retrieval is separated from recommendation ranking, and authentication is separated from user settings and item management.

\section{Implementation}
\subsection{Core Data and CRUD}
To satisfy the coursework requirement for database-backed CRUD, the \texttt{Item} model implements full RESTful CRUD:
\begin{itemize}
  \item \texttt{POST /api/items} creates an item,
  \item \texttt{GET /api/items} and \texttt{GET /api/items/:id} read user items,
  \item \texttt{PATCH /api/items/:id} updates an item,
  \item \texttt{DELETE /api/items/:id} deletes an item.
\end{itemize}
These operations are backed by PostgreSQL through Prisma and return conventional JSON/status-code responses.

The \texttt{Item} model was the most appropriate candidate for full CRUD because it represents a stable user-owned resource with clear lifecycle operations.
The listings subsystem was also extended to support a fuller lifecycle: listings can now be updated or deleted by their owner, and deleting a listing releases the linked item back to the \texttt{AVAILABLE} state instead of leaving it trapped in a reserved marketplace state.

\subsection{Recommendation Pipeline}
The recommendation service combines three signals. First, climate suitability measures whether an outfit is likely to be comfortable for the chosen location and time. Second, trend relevance applies a bounded boost or penalty from recent media-derived signals. Third, user feedback introduces a small adaptive bias so the system becomes more personalised without becoming unstable. The recommendation endpoint remains publicly usable, but it now supports optional authenticated requests so a valid JWT-bearing session can pass per-user context into scoring. I deliberately kept this model heuristic and bounded: it is easier to explain, debug, and defend than an opaque ranking method, and it fits a client flow where users can search from a curated city list, enter coordinates directly, or reuse a saved home location.

This design is especially useful in the target use case. A user getting ready quickly may know that the day is ``cold'' or ``rainy'', but not whether the weather will change by the time they leave. By consolidating these signals into one recommendation, the system reduces fragmented decision-making and supports comfort and self-presentation.

The recommendation flow is therefore also an information-fusion problem. Weather APIs supply environmental context, trend processing adds cultural context, and user history contributes personal context. Open-Meteo was particularly useful because it offers forecast and geocoding support through a lightweight interface \cite{openmeteo}.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  node distance=5mm and 6mm,
  every node/.style={font=\footnotesize},
  box/.style={draw, rounded corners, align=center, text width=0.2\linewidth, minimum height=0.7cm, inner sep=3pt},
  wide/.style={draw, rounded corners, align=center, text width=0.24\linewidth, minimum height=0.7cm, inner sep=3pt},
  arr/.style={-{Latex[length=2mm]}, thick}
]
\node[box] (weather) {Weather context};
\node[box, right=of weather] (trend) {Trend signals};
\node[box, right=of trend] (user) {User history};
\node[wide, below=of trend] (score) {Scoring and ranking};
\node[wide, below=of score] (output) {Recommendation + explanation};
\draw[arr] (weather) -- (score);
\draw[arr] (trend) -- (score);
\draw[arr] (user) -- (score);
\draw[arr] (score) -- (output);
\end{tikzpicture}
\caption{Recommendation data flow.}
\label{fig:recflow}
\end{figure}

\subsection{Data Flow Through the System}
The operational flow begins when the user authenticates and sends wardrobe data or a recommendation request. The client sends structured JSON payloads to the API, where validation and access checks run before business logic. A saved account location can prefill the recommendation form, while blank settings intentionally do not force an autofill. The recommendation service then combines internal state from PostgreSQL with external weather and trend context. After scoring candidate outfits, the API returns a JSON response containing the selected result, and logged-in users can store that recommendation with its weather, place, and time for later review.

\subsection{Security Decisions}
Security was addressed at multiple layers.

\textbf{Authentication.} The API uses JWT-based authentication with guarded routes. This provides a simple stateless security model for a coursework API and allows protected ownership checks without server-side session storage \cite{jwt}. The account area also supports authenticated password change with current-password verification before a new bcrypt hash is stored. JWT secrets are treated as required environment configuration rather than silently falling back to hard-coded defaults.

\textbf{Authorisation.} Ownership checks were added to CRUD operations and saved recommendations so a user can only read, update, or delete their own records. This prevents horizontal privilege issues and ensures the API does not expose another user's wardrobe data.

\textbf{Validation.} DTO validation is enforced in NestJS using class-validator and a global validation pipe. This helps prevent malformed payloads from entering business logic or the database layer, including password-policy enforcement and structured validation of saved account location input.

\textbf{Secret handling.} Environment files are excluded from version control, and previously tracked \texttt{.env} data was removed from Git history. This is important because leaked API keys invalidate any claim of basic operational security.

\textbf{Proxy policy.} The image proxy used for Pinterest inspiration was tightened so it only accepts HTTPS requests to approved Pinterest image hosts, applies a timeout, and verifies that the upstream response is an image. This reduces the risk of turning a convenience proxy into a general server-side fetch endpoint.

\textbf{Database safety.} Prisma reduces query-construction mistakes by using typed queries rather than manual string assembly, lowering the risk of common implementation errors.

\textbf{Error handling.} Conventional HTTP status codes are part of the security and reliability model. Explicit \texttt{401}, \texttt{403}, and \texttt{404} responses make access failures visible and predictable instead of ambiguous. This is particularly important for ownership-controlled CRUD endpoints, where incorrect status behaviour can easily hide security mistakes during testing.

\textbf{Operational trade-off.} JWT was selected because it is simple and stateless, but token revocation is weaker than in a full server-side session model. For this coursework-scale system, simplicity and ease of integration outweighed that limitation.

\section{Evaluation}
\subsection{Usefulness}
The system is useful because it reduces cognitive load during a time-constrained daily task. Users may know the current weather in general terms but still miss details such as rain probability, wind, or conditions at a later time of day. They may also have only vague awareness of recent fashion trends. Style Forecast aggregates those signals and turns them into a single actionable suggestion, reducing both uncertainty and decision fatigue.

\subsection{Testing and Validation}
Evaluation was carried out in four ways:
\begin{enumerate}
  \item backend and frontend production builds to confirm deployable integrity,
  \item endpoint validation through Swagger and direct HTTP requests,
  \item automated unit tests for recommendation and account logic,
  \item end-to-end flow checks across authentication, item creation, listing lifecycle, recommendation, saving, and feedback.
\end{enumerate}

This evaluation strategy reflects how the system is actually consumed. A successful build confirms structural integrity, but the more important question is whether a realistic user flow behaves correctly from login through to recommendation output.

The evaluation is qualitative rather than benchmark-driven. The main question is whether the recommender gives plausible, explainable, and secure recommendations under realistic usage conditions.

\subsection{Challenges and Lessons}
The most difficult implementation issue was timezone correctness. Weather relevance depends on when and where the user intends to wear an outfit, so naive datetime handling produces incorrect recommendations. This also affected account settings, because the interface needed to display an obvious user-facing timezone label while still storing a robust underlying zone identifier. A second challenge was external data reliability, so the system had to fail clearly rather than silently.

The main lesson from the project was that explainability matters as much as raw feature count. A smaller system with clear logic, secure defaults, and a strong end-to-end flow is more defensible than a larger but less coherent implementation.

Another lesson was architectural discipline. Separating concerns between controller, service, and persistence layers added some overhead early on, but it paid off when the CRUD requirement, JWT refresh flow, and documentation changes had to be introduced without destabilising unrelated behaviour.

A further lesson was that external intelligence should be bounded. Trend and weather integrations can improve usefulness, but they should not dominate core recommendation logic.

\subsection{Limitations and Future Work}
The current trend sentiment model is lexical and can miss nuance. Personalisation is heuristic rather than learned from large-scale data, and automated test coverage remains narrower than production standards. A scheduled email-digest channel was explored, but reliable SMTP or OAuth-backed delivery introduced provider and deployment dependencies that were disproportionate to the coursework scope. Future work would therefore focus on stronger automated testing, scheduled trend-ingestion jobs, and a constrained learning-to-rank approach that preserves explainability.

\section{Conclusion}
Style Forecast was designed to bridge a genuine everyday knowledge gap: what to wear when time is limited and contextual information is fragmented. The final architecture reflects deliberate choices about maintainability, security, explainability, and practical usefulness.

\paragraph*{Generative AI Declaration}
Generative AI tools were used selectively for ideation and drafting assistance. Final design choices and report wording were reviewed and finalised by me.

\begingroup
\footnotesize
\setlength{\itemsep}{0pt}
\setlength{\parskip}{0pt}
\setlength{\topsep}{2pt}
\setlength{\partopsep}{0pt}
\setlength{\parsep}{0pt}
\raggedright
\begin{thebibliography}{9}
\bibitem{typescript} Microsoft. \emph{TypeScript Docs}. \url{https://www.typescriptlang.org/}.
\bibitem{nestjs} NestJS. \emph{NestJS Docs}. \url{https://docs.nestjs.com/}.
\bibitem{prisma} Prisma. \emph{Prisma ORM Docs}. \url{https://www.prisma.io/docs}.
\bibitem{postgresql} PostgreSQL Global Development Group. \emph{PostgreSQL Docs}. \url{https://www.postgresql.org/docs/}.
\bibitem{openapi} OpenAPI Initiative. \emph{OpenAPI Specification}. \url{https://spec.openapis.org/oas/latest.html}.
\bibitem{react} Meta. \emph{React Docs}. \url{https://react.dev/}.
\bibitem{vite} Vite. \emph{Vite Guide}. \url{https://vite.dev/guide/}.
\bibitem{openmeteo} Open-Meteo. \emph{Weather Forecast API Docs}. \url{https://open-meteo.com/en/docs}.
\bibitem{jwt} Jones, Bradley, and Sakimura. \emph{RFC 7519: JSON Web Token (JWT)}. \url{https://www.rfc-editor.org/rfc/rfc7519}.
\end{thebibliography}
\endgroup

\end{document}
